Score-based tests of measurement invariance: Use in practice

### 2013-09-15 frontiers/frontiers.Rnw
organize cat2bb, organize path diagram, organize general discussion, organize sim2 results
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@259 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@

\fixme{Need to update the citation below so that the last name is ``van de Schoot''.}
\seeme{Is it a different paper from vande12?}
Measurement invariance is most often tested using multiple groups models \cite<see, e.g., >{vande12}.  
This amounts to assuming that our auxiliary variable is categorical (i.e., unordered), 
which is not true for the age groups in the data.  However, we demonstrate the procedure for completeness.

@@

## Unrestricted model
full <- cfa('f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5',
             data=yg, group="agegroup", meanstructure=TRUE)
@
Finally, we can get the results of a likelihood ratio test via the \verb+anova()+ 
function, which implies that the GQ-6 violates measurement invariance.
<<LRT>>=
## LRT
anova(full, reduced)
@
Because our sample size is large, this likelihood ratio test is sensitive to 
small measurement invariance violations \cite{Ben80}.  Thus, the above test may 
be significant solely due to the large sample size.  Additionally, the above test 
is sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  We need test statistics that 
are sensitive to measurement invariance violations that are monotonic 
with age.  The score-based statistics described previously possess this property.

%% TODO Illustrate catL2BB statistic?

@@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM")

round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
In the code above, the \verb+parm=1:4+ argument means that 
we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
which are supplied by lavaan.  The ordering of parameters can be obtained via the 
command \verb+coef(reduced)+.  We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

<<ordtest,eval=FALSE>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo")
@ 
<<pvals>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}[H]
\caption{Fluctuation processes for the WDM statistic(left panel) and the maxLM statistic(right panel)}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE>>=
par(mfrow = c(1, 2), cex = 0.65)
WDMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="WDMo", fplot=TRUE)
#WDMo 
maxLMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="maxLMo", fnl=opval1, fplot=TRUE)
#maxLMo 
@
\end{figure}

%% TODO Describe graphs
%% Is there some argument for fplot=TRUE so user can change  axis-name and critical value line
Plots representing the statistics' fluctuations across levels of age group are displayed in Figure~\ref{fig:ordres}. The left panel displays the process associated with WDM. The right panel displays the process associated with maxLM. In both cases, the test statistics in the sequence assess a split of the observations up to age group $i$ vs. greater than $i$, and the null hypothesis is rejected if the maximum of the statistics is larger than its 5\% critical value(visualized by the horizontal red line). The final group is not displayed because it encompass all observations and always equals to zero. It is observed that both statistics generally increase with age. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Power and Type I error were examined across three sample sizes
($n=120,480,960$) , three numbers of categories ($m=4,8,12$)and
17 magnitudes of invariance violations. 
The measurement invariance violations began at level $1+m/2$ of $V$
and were consistent thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested. Four statistics were examined above. In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("ordmax","ordwmax","catdiff")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("ordmax","ordwmax","catdiff"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Loading1.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("loading"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Covariance}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("var"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Error1}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("error"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra, unmodeled loading. 
Sample size and magnitude of measurementinvariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

\subsection{Results}

%<<sim2>>=
%## seed for replication
%RNGkind(kind = "default", normal.kind = "default")
%set.seed(1163)

%if(file.exists("sim2.rda")) {
%  load("sim2.rda")
%} else {
%  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))
%  sim2$nlevels <- factor(sim2$nlevels)
%  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
%  sim2$nobs <- factor(sim2$nobs)
%  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
%  levels(sim2$test) <- c("ordmax","ordwmax","catdiff")
%  save(sim2, file="sim2.rda")
%}
%sim2$test <- factor(as.character(sim2$test),
%  levels = c("ordmax","ordwmax","catdiff"),
%  labels = c("ordmax","ordwmax","catdiff"))
%@

@@

The core of the proposed ordinal test is to get the cumulative score process, which can be applied to any ML estimation functions or functions relating to maximizing or minimizing some statistics quantity.

>>>>

%% TODO Illustrate catL2BB statistic

If $V$ is only nominal/categorical,
there is not even a partial ordering, i.e., measurement invariance tests should neither
exploit the ordering of $V$'s levels nor of the observations within the level.
In this situation, unordered $\mathit{LM}$ is proposed.  This test statistic can be formally written as
\begin{equation}
    \label{eq:lmuo}
     \mathit{LM}_\mathit{uo} = \sum_{\ell = 1, \dots, m} \sum_{j = 1, \dots, k}
       \left( {\bm B}(\hat {\bm \theta})_{i_\ell j} - {\bm B}(\hat {\bm \theta})_{i_{\ell - 1}j} \right)^2,
\end{equation}
This test statistic discards the ordinal nature of the auxiliary
variable, essentially similar to the nature of likelihood ratio tests (or,
equivalently, via Wald tests or Lagrange multiplier tests): the ordinality of the auxiliary variable is
ignored. By contrast, this statistics is easier to get compared to likelihood ratio test statistics since it doesn't involve specifying different sets of competitive models.

@@

Measurement invariance is most often tested using multiple groups models \cite<see, e.g., >{vande12}.  
This amounts to assuming that our auxiliary variable is categorical (i.e., unordered), 
which is not true for the age groups in the data.  However, we demonstrate the procedure for completeness.

@@

## Unrestricted model
full <- cfa('f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5',
             data=yg, group="agegroup", meanstructure=TRUE)
@
Finally, we can get the results of a likelihood ratio test via the \verb+anova()+ 
function, which implies that the GQ-6 violates measurement invariance.
<<LRT>>=
## LRT
anova(full, reduced)
@

Another similar statistics is $\mathit{LM}_\mathit{uo}$ as shown in Equation~ \eqref{eq:lmuo}. We can obtain this statistic by using the following function. 
<<LM_UO>>=
## unordered LM
lm_u0 <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="range")
round(c(lm_u0$p.value), 2)
@
we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
which are supplied by lavaan.  The ordering of parameters can be obtained via the 
command \verb+coef(reduced)+.

Because our sample size is large, this likelihood ratio test is sensitive to 
small measurement invariance violations \cite{Ben80}.  Thus, the above test may 
be significant solely due to the large sample size.  Additionally, the above test 
is sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  We need test statistics that 
are sensitive to measurement invariance violations that are monotonic 
with age.  The score-based statistics described previously possess this property.

@@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM")
round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
%% move this code description to the cat2BB part. 
%In the code above, the \verb+parm=1:4+ argument means that 
%we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
%which are supplied by lavaan.  The ordering of parameters can be obtained via the 
%command \verb+coef(reduced)+.

We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

<<ordtest,eval=FALSE>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo")

@ 
<<pvals>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@

Besides, plots can be generated by using \verb+fplot=TRUE+ argument inside of \verb+sctest.lavaan()+ function. Plots representing the statistics' fluctuations across levels of age group are displayed in Figure~\ref{fig:ordres}. The left panel displays the process associated with $\mathit{WDM}$. The right panel displays the process associated with $\mathit{maxLM}$. In both cases, the test statistics in the sequence assess a split of the observations up to age group $i$ vs. greater than $i$, and the null hypothesis is rejected if the maximum of the statistics is larger than its 5\% critical value(visualized by the horizontal red line). The final group is not displayed because it encompass all observations and always equals to zero. The age group 1:5 in the x-axis corresponds to 10-11, 12-13, 14,15,16 respectively. It is observed that both statistics generally increase with age.

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}[H]
\caption{Fluctuation processes for the WDM statistic(left panel) and the maxLM statistic(right panel)}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE>>=
par(mfrow = c(1, 2), cex = 0.65)
WDMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="WDMo", fplot=TRUE,ylim=c(0,3.5),xlab="Age group",ylab = "Weighted max statistics")

#WDMo 
maxLMo<-sctest(reduced,as.numeric(yg$agegroup),parm=1:4, stat="maxLMo", fnl=opval1, fplot=TRUE,ylim=c(0,14),xlab="Age group",ylab = "LM statistics")
#maxLMo 
@
\end{figure}

\seeme{can't change age group axis, since order.by argument needs agegroup as numeric}

%% TODO Describe graphs
%% Is there some argument for fplot=TRUE so user can change  axis-name and critical value line

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[H]
\caption{Path diagram representing the factor analysis model 
         used for simulations.}
\label{fig:path}
\includegraphics{PathDiagram}
\end{figure}

Power and Type I error were examined across three sample sizes
($n=120,480,960$), three numbers of categories ($m=4,8,12$) and
17 magnitudes of invariance violations. 
The measurement invariance violations began at level $1+m/2$ of $V$
and were consistent thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested. Four statistics were examined above. In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

@@

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Loading1.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$parms %in% c("loading")& sim1$pars %in% c("Loading1","All Loadings","All Errors"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Covariance}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("maxLM_o","WDM_o","LM_uo")&sim1$parms %in% c("var") &sim1$pars %in% c("All Loadings", "Covariance","All Errors") ,]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Error1}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("maxLM_o","WDM_o","LM_uo")&sim1$parms %in% c("error") &sim1$pars %in% c("Error1","All Errors","All Loadings"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra, unmodeled loading. 
Sample size and magnitude of measurement invariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

\subsection{Results}
Results are presented by the parameter that violated measurement invariance.
Figure~\ref{fig:sim21res} displays results when the unmodeled loading violates 
measurement invariance. One can generally observe that 
the first loading and error variance parameter showed high measurement invariance 
violation simultaneously, while factor covariance did not demonstrate violation.

When the loading $\lambda_{11}$ violates
measurement invariance, even though lacking one loading associated with the first indicator, 
statistic showed enough power to detect the first loading violation specifically. 
Similar pattern of specificity and power were observed when the violation parameter 
was $\sigma_{12}$ and $\epsilon_{11}$. Figures are shown in Appendix for simplicity.

In summary, the proposed test statistics appear robust to unmodeled loading parameters,so long
as the umodeled parameter does not violate measurment invariance. If the unmodeled parameter 
dose violate measurement invariance, however, then the tests assign this violation to other 
parameters that do not violate measurement invariance. The impacted parameters include the error
variance and other loadings associated with the manifest variables that had the unmodeled 
loading. Thus, prior to testing measurement invariance, it is important to study the extent to
which the hypothesized model includes all parameters of importance.

<<sim2>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))

  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
  levels(sim2$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim21res}
\setkeys{Gin}{width=\textwidth}
<<sim21res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("maxLM_o","WDM_o","LM_uo")&sim2$parms %in% c("extra")&sim2$pars %in% c("Loading1","Error1","Covariance"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

We focused on testing for measurement invariance in factor analysis model in this paper, however, the core of the proposed ordinal test is to get the cumulative score process, which can be applied to any ML estimation functions or functions relating to maximizing or minimizing some statistics quantity. For example, studying DIF in IRT is immediate.

In summary, the ordinal test statistic proposed have relatively-high power for detecting measurement invariance violation parameter specifically that are monotonic with the auxiliary ordinal variable. This
feature is robust as long as the violation parameter is included in the testing model. Furthermore, as illustrated in tutorial part, the use of these tests is convenient without changing the focal psychometric model. In all, the tests have advantageous properties that should be implemented in practice.

\section{Appendix}
\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim22res}
\setkeys{Gin}{width=\textwidth}
<<sim22res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("maxLM_o","WDM_o","LM_uo")&sim2$parms %in% c("extra+loading"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim23res}
\setkeys{Gin}{width=\textwidth}
<<sim23res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("maxLM_o","WDM_o","LM_uo")&sim2$parms %in% c("extra+var"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim24res}
\setkeys{Gin}{width=\textwidth}
<<sim24res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("maxLM_o","WDM_o","LM_uo")&sim2$parms %in% c("extra+error"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@
